The Civil Society Organizations effect: A mixed-Methods analysis of bottom-up approaches in Brazilian public policy.

### 2023-11-06 Untitled/Untitled.Rmd
paper R&P

---
title: "The Civil Society Organizations effect: A mixed-Methods analysis of bottom-up approaches in Brazilian public policy"
date: "2023-10-26"
output: bookdown::pdf_document2
sansfont: Sans Serif
fontsize: 12pt
author:
  - name: Manoel Galdino
    affiliation: Universidade de São Paulo - Brazil.
  - names: Juliana Sakai
    affiliation: Transparência Brasil - Brazil
  - names: Bianca Vaz Mondo
    affiliation: Transparência Brasil - Brazil
  - names: Natália Paiva
    affiliation: Alandar Consulting - Brazil
abstract: "This paper examines the effect of bottom-up accountability on public service delivery. We differentiate between information-driven interventions and the mobilization and monitoring efforts of organized Civil Society Organizations (CSOs), and argue that the latter type of interventions can can drive significant policy change. The study evaluates the effectiveness of the Obra Transparente project by Brazilian NGO Transparência Brasil, engaging 21 local CSOs in South and Southeast Brazil.  Our findings emphasize the importance of sustained, coordinated efforts by socially embedded CSOS. These efforts, involving direct engagement with municipal officials and ensuring that their complains cannot be ignored lead to more substantial outcomes when compared with citizens information-driven."
keywords: "accountability, Civil Society Organizations, mixed methods, community-driven development, public policy."
bibliography: OT_paper.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

## Introduction

Accountability is a crucial element in democratic systems, ensuring the efficient delivery of public services [@besley_2003; @cameron_2004; @odonnell_1998]. Initially defined as the process of holding authorities responsible for their actions [@mulgan_2000; @oloughlin_1990], the literature has since broadened to include various oversight mechanisms for politicians and the bureaucracy. Some authors focus on formal institutions [@odonnell_2003; @kenney_2003], known as horizontal accountability, while others examine how politicians respond to voters, the media, and civil society organizations [@bertelli_2020; @relly_2012; @paul_1992], referred to as vertical accountability.

@@

The evidence gathered is primarily based on documented reports collected by Transparência Brasil and produced by the local OSBs over the course of the project. We also use evidence collected on semi-structured interviews and an online survey with representatives of the 21 CSOs conducted by an independent evaluator of the project. For the quantitative analysis, we use administrative data about construction works data by municipalities and socio-demographic data on municipalities. The questions asked in the survey as well the aggregate answered are presented in annex A.

@@

# these are the no redundancy covariates with and without state covariate
covs.nr <- c('perc_pea18', 'ext_pobres', 'log_rpc', 'p10_ricos_desigualdade', 'freq_escola_15_1', 'analfabetismo',
             'log_renda', 'log_pop', 'percentual_de_execucao', 'tipo_do_projeto')

covs.nr.st <- c(covs.nr, "uf.x")

>>>>

---
title: "The Civil Society Organizations effect: A mixed-Methods analysis of bottom-up approaches in Brazilian public policy."
short: "The Civil Society Organizations effect"
month: "`r format(Sys.Date(), '%m')`"
year: "`r format(Sys.Date(), '%Y')`"
keywords:
  - accountability
  - Civil Society Organizations
  - mixed methods
  - community-driven development
  - public policy.
author:
  - name: Manoel Galdino
    firstname: Manoel
    surname: Galdino
    email: mgaldino@usp.br
    affiliation: Universidade de São Paulo, Brazil
  - name: Bianca Vaz Mondo
    firstname: Bianca
    surname: Mondo
    affiliation: Government Transparency Institute - Germany
  - name: Juliana Mari Sakai
    firstname: Juliana
    surname: Sakai
    affiliation: Transparência Brasil - Brazil
  - name: Natália Paiva
    firstname: Natália
    surname: Paiva
    affiliation: Alandar Consulting - Germany
acknowledgements: |
  We would like to thank all participants of project Obra Transparente, Vanda Medeiros and the team of Transparência Brasil who worked in the project or in evaluations of the project. The project was funded by an UNDEF grant, which we also thank.
abstract: |
  This paper examines the effect of bottom-up accountability on public service delivery. We differentiate between information-driven interventions and the mobilization and monitoring efforts of organized Civil Society Organizations (CSOs), and argue that the latter type of interventions can can drive significant policy change. The study evaluates the effectiveness of the Obra Transparente project by Brazilian NGO Transparência Brasil, engaging 21 local CSOs in South and Southeast Brazil.  Our findings emphasize the importance of sustained, coordinated efforts by socially embedded CSOS. These efforts, involving direct engagement with municipal officials and ensuring that their complains cannot be ignored lead to more substantial outcomes when compared with citizens information-driven.
bibliography: OT_paper.bib
output: rticles::aea_article
---

@@

## Introduction
Accountability is a crucial element in democratic systems, ensuring the efficient delivery of public services [@besley_2003; @cameron_2004; @odonnell_1998]. Initially defined as the process of holding authorities responsible for their actions [@mulgan_2000; @oloughlin_1990], the literature has since broadened to include various oversight mechanisms for politicians and the bureaucracy. Some authors focus on formal institutions [@odonnell_2003; @kenney_2003], known as horizontal accountability, while others examine how politicians respond to voters, the media, and civil society organizations [@bertelli_2020; @relly_2012; @paul_1992], referred to as vertical accountability.

@@

The evidence gathered is primarily based on documented reports collected by Transparência Brasil and produced by the local OSBs over the course of the project. We also use evidence collected on semi-structured interviews and an online survey with representatives of the 21 CSOs conducted by an independent evaluator of the project. For the quantitative analysis, we use administrative data about construction works data by municipalities and socio-demographic data on municipalities. The questions asked in the survey as well the aggregate answered are presented in annex B.

@@
